Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-08-12 manuscript/main.tex
in the middle of the orthogonality proof

\subsection{Preliminary step 1: finite-basis approximation}\label{sec: finite basis approximation}	
Let ${b_m(t)}$ be a twice-differentiable basis of $L^2([0,1])$ whose second derivatives are also linearly independent, for example, cubic B-splines, the Fourier basis, or an orthonormal polynomial basis of degree greater than three. Using this basis, 
the $j$th functional predictor, 
regression coefficient, PLS component direction, and orthogonalization regression coefficient (with iteration indices suppressed) are represented as follows:
\begin{equation*}
	X_{ij}(t) = \sum_{m=1}^\infty \theta_{ijm} \, b_m(t),~
	\beta_j(t) = \sum_{m=1}^\infty \eta_{jm} \, b_m(t),~
	%%%%%%%%%%%%%%%%%%%%%%
	\xi_j(t)
	=
	\sum_{m=1}^\infty
	\gamma_{jm}
	b_m(t),
	~
	\delta_j(t)
	=
	\sum_{m=1}^\infty
	\pi_{jm}
	b_m(t)		
\end{equation*}	
In practice, the full set of coefficients can not be obtained  with finite sample size, as functional data are measured on a finite grid. Thus, we truncate the expansion at $M$ terms and define the truncated basis space
$
\mathcal{B} := \operatorname{span}\bigl(b_1(t), \ldots, b_M(t)\bigr).
$
We choose a moderately large $M$ (e.g., 15 or 20) to capture functional variation without fine-tuning, as smoothness is handled via penalization (see Section~\ref{sec: Regularization of PLS Components}).
The truncated expansions of the predictor and coefficient are denoted as
\begin{equation*}
	\widetilde{X}_{ij}(t) := \sum_{m=1}^M \theta_{ijm} \, b_m(t), \quad
	\tilde{\beta}_j(t) := \sum_{m=1}^M \eta_{jm} \, b_m(t),
\end{equation*}
and our suggest method restricts each PLS component direction and orthogonalization regression coefficient to admit the following expansion:
\begin{equation}\label{def:truncated_zeta_and_delta}
	\xi_j(t) = \sum_{m=1}^M \gamma_{jm} \, b_m(t),
	\quad
	\delta_j(t)
	=
	\sum_{m=1}^M
	\pi_{jm}
	b_m(t)		
\end{equation}
This implies that all computations in this paper are carried out entirely within the subspace
\begin{equation}\label{def:truncated_hilbert_space}
	\widetilde{ \mathcal{H} } := \mathcal{B}^K \times \mathbb{R}^p \subset \mathcal{H}.
\end{equation}
The $i$th hybrid predictor, projected on $\widetilde{ \mathcal{H} }$, is represented by the tuple
\begin{equation*}
	\widetilde{W}_i := 
	(
	\widetilde{X}_{i1}, \ldots, \widetilde{X}_{iK}, \boldsymbol{Z}_i
	).
\end{equation*}
Let $\boldsymbol{\theta}_{ij}$, $\boldsymbol{\eta}_j$,  $\boldsymbol{\gamma}_j$, and
$\boldsymbol{\pi}_j$
denote the $M$-dimensional vectors of coefficients:
\begin{equation}\label{def:coef_vector_form}
	\boldsymbol{\theta}_{ij} := (\theta_{ij1}, \ldots, \theta_{ijM})^\top,
	~
	\boldsymbol{\eta}_{j} := (\eta_{j1}, \ldots, \eta_{jM})^\top ,~
	\boldsymbol{\gamma}_{j} := (\gamma_{j1}, \ldots, \gamma_{jM})^\top,~
	\boldsymbol{\pi}_{j} := (\pi_{j1}, \ldots, \pi_{jM})^\top
\end{equation}
For the predictors, we stack the coefficient vectors across observations into the matrix
\begin{equation}\label{def:theta}
	\Theta_j := (\boldsymbol{\theta}_{1j}, \ldots, \boldsymbol{\theta}_{nj})^\top  \in \mathbb{R}^{n \times M},
\end{equation}
and construct the full design matrix $\Theta := (\Theta_1, \ldots, \Theta_K, Z) \in \mathbb{R}^{n \times (MK + p)}$. Let us denote the response vector as $\mathbf{y} := (y_1, \ldots, y_n)^\top$.
Let $B, B^{\prime\prime} \in \mathbb{R}^{M \times M}$ denote the Gram matrices of the basis functions and their second derivatives, with entries
\begin{equation}\label{def:B_matrix}
	B_{m, m'} := \int_0^1 b_m(t)\, b_{m'}(t)\, dt, \quad
	B^{\prime\prime}_{m, m'} := \int_0^1 b_m''(t)\, b_{m'}''(t)\, dt,
\end{equation}
for $m, m' = 1, \ldots, M$. We then define the block-diagonal matrices
\begin{equation}\label{def:gram_block}
	\mathbb{B} := \operatorname{blkdiag}(B, \ldots, B, I_p), \quad
	\mathbb{B}^{\prime\prime} := \operatorname{blkdiag}(B^{\prime\prime}, \ldots, B^{\prime\prime}, I_p),
\end{equation}

@@

		\langle
		\xi_j, \xi_j
		\rangle

@@

\section{Proof of Lemma \ref{proposition:closed_form_orthgonalization} }\label{section:proof:proposition:closed_form_orthgonalization}
For notational simplicity, we omit the iteration superscript $[l]$.
Using the basis expansion coefficient notation from \eqref{def:truncated_zeta_and_delta}, \eqref{def:coef_vector_form}, \eqref{def:theta}, and \eqref{def:B_matrix}, the full vector of PLS scores $\hat{\boldsymbol{\rho}}  = (\hat{\rho}_1, \ldots, \hat{\rho}_n)^\top$ 
where
$\hat{\rho}_1$ is defined in \eqref{def:plsscore},
is computed through the following matrix multiplication:
\begin{equation*}
	\hat{\boldsymbol{\rho}}^\top = \sum_{k=1}^K ( \hat{\boldsymbol{\gamma}}_j )^\top B \, \Theta_j^\top + \boldsymbol{\zeta}^\top \mathbf{Z}^\top.
\end{equation*}
The criterion $\operatorname{PENSSE}( \delta)$ can be decomposed as follows:

@@

Firstly, it is obvious by the unit-norm constraint enforced in Proposition \ref{proposition: eigenanalysis} that:
\begin{align*}
	\langle \langle \hat{\xi}_l,  \hat{\xi}_l \rangle \rangle &= \langle \widehat{\xi}_l, \widehat{\xi}_j \rangle_{\mathcal{H}} + \sum_{k=1}^K \lambda_k \left\langle \ddot{\widehat{\psi}}_l^{(k)}, \ddot{\widehat{\psi}}_j^{(k)} \right\rangle_{L^2}  \\
	&= \sum \limits_{k=1}^K \int_0^1 \widehat{\psi}_l^{(k)}(t_k) \widehat{\psi}_j^{(k)}(t_k) dt_k + \sum \limits_{r=1}^p \widehat{\theta}_{lr} \widehat{\theta}_{jr} + \sum_{k=1}^K \lambda_k \int_0^1 \ddot{\widehat{\psi}}_l^{(k)}(t_k) \ddot{\widehat{\psi}}_j^{(k)}(t_k) dt_k \\
	&= \sum \limits_{k=1}^K \int_0^1 \hat{\mathbf{a}}^{(k) \top}_l \mathbf{b}^{(k)}(t_k) \mathbf{b}^{(k) \top}(t_k) \hat{\mathbf{a}}^{(k)}_l  dt_k + \boldsymbol{\theta}_l^\top \boldsymbol{\theta}_l + \sum \limits_{k=1}^K \lambda_k \int_0^1 \hat{\mathbf{a}}^{(k) \top}_l \ddot{\mathbf{b}}^{(k)}(t_k) \ddot{\mathbf{b}}^{(k) \top}(t_k) \hat{\mathbf{a}}^{(k)}_l  dt_k \\
	& = \sum \limits_{k=1}^K \hat{\mathbf{a}}_l^{(k) \top} J^{(k)} \hat{\mathbf{a}}_l^{(k)} + \boldsymbol{\theta}_l^\top \boldsymbol{\theta}_l +  \sum \limits_{k=1}^K \lambda_k \hat{\mathbf{a}}^{(k) \top}_l \ddot{J}^{(k)} \hat{\mathbf{a}}^{(k)}_l \\
	& = \hat{\xi}_l^{*\top} J^\ast \hat{\xi}_l^* + \hat{\xi}_l^{*\top} \Lambda \ddot{J}^\ast \hat{\xi}_l^* = \hat{\xi}_l^{*\top} (J^\ast +\Lambda \ddot{J}^\ast) \hat{\xi}_l^* = 1
\end{align*}
where $\ddot{\mathbf{b}}^{(k)}(t_k) = (\ddot{b}_1^{(k)}(t_k), \ldots, \ddot{b}_M^{(k)}(t_k))^\top$

Now to prove the orthogonality, we first express the $l$-th residualized hybrid predictor in terms of its previous $(l-1)$-th version:
	& =\widetilde{C}^{*[l-1]}B^*[\mathbf{t}] -  \left(\frac{\widehat{\boldsymbol{\rho}}_{l-1} \widehat{\boldsymbol{\rho}}_{l-1}^\top}{\widehat{\boldsymbol{\rho}}_{l-1}^\top \widehat{\boldsymbol{\rho}}_{l-1}} \right) \widetilde{C}^{*[l-1]} B^*[\mathbf{t}] \\
	&= \left(I_n - \frac{\widehat{\boldsymbol{\rho}}_{l-1} \widehat{\boldsymbol{\rho}}_{l-1}^\top}{\widehat{\boldsymbol{\rho}}_{l-1}^\top \widehat{\boldsymbol{\rho}}_{l-1}} \right) \widetilde{C}^{*[l-1]} B^*[\mathbf{t}].
Thus, we have
	\widetilde{W}^{[l]}[\mathbf{t}] &= \left(I_n - \frac{\widehat{\boldsymbol{\rho}}_{l-1} \widehat{\boldsymbol{\rho}}_{l-1}^\top}{\widehat{\boldsymbol{\rho}}_{l-1}^\top \widehat{\boldsymbol{\rho}}_{l-1}} \right) \widetilde{W}^{[l-1]}[\mathbf{t}] \\
	&= \left(I_n - \frac{\widehat{\boldsymbol{\rho}}_{l-1} \widehat{\boldsymbol{\rho}}_{l-1}^\top}{\widehat{\boldsymbol{\rho}}_{l-1}^\top \widehat{\boldsymbol{\rho}}_{l-1}} \right) \left(I_n - \frac{\widehat{\boldsymbol{\rho}}_{l-2} \widehat{\boldsymbol{\rho}}_{l-2}^\top}{\widehat{\boldsymbol{\rho}}_{l-2}^\top \widehat{\boldsymbol{\rho}}_{l-2}} \right) \widetilde{W}^{[l-2]}[\mathbf{t}] \\
	&= P \left(I_n - \frac{\widehat{\boldsymbol{\rho}}_{j} \widehat{\boldsymbol{\rho}}_{j}^\top}{\widehat{\boldsymbol{\rho}}_{j}^\top \widehat{\boldsymbol{\rho}}_{j}} \right)\widetilde{W}^{[j]}[\mathbf{t}]
where $P$ is an $n \times n$ matrix defined as
	P = \prod \limits_{h=j+1}^{l-1} \left(I_n - \frac{\widehat{\boldsymbol{\rho}}_{h} \widehat{\boldsymbol{\rho}}_{h}^\top}{\widehat{\boldsymbol{\rho}}_{h}^\top \widehat{\boldsymbol{\rho}}_{h}} \right)
Then for $j < l$,
	\int_{\mathcal{T}} \widetilde{W}^{[l]}[\mathbf{t}] \widehat{\xi}_j[\mathbf{t}] d\mathbf{t} &= \int_{\mathcal{T}} P   \left(I_n - \frac{\widehat{\boldsymbol{\rho}}_{j} \widehat{\boldsymbol{\rho}}_{j}^\top}{\widehat{\boldsymbol{\rho}}_{j}^\top \widehat{\boldsymbol{\rho}}_{j}} \right)\widetilde{W}^{[j]}[\mathbf{t}]  \widehat{\xi}_k[\mathbf{t}] d\mathbf{t}  \\
	&= P \left\{\int_{\mathcal{T}} \widetilde{W}^{[j]}[\mathbf{t}] \widehat{\xi}_j[\mathbf{t}] d\mathbf{t} - \frac{\widehat{\boldsymbol{\rho}}_{j} \widehat{\boldsymbol{\rho}}_{j}^\top}{\widehat{\boldsymbol{\rho}}_{j}^\top \widehat{\boldsymbol{\rho}}_{j}} \int_{\mathcal{T}} \widetilde{W}^{[j]}[\mathbf{t}] \widehat{\xi}_j[\mathbf{t}] d\mathbf{t} \right\} \\ 
	&=P \left( \widehat{\boldsymbol{\rho}}_j - \frac{\widehat{\boldsymbol{\rho}}_{j} \widehat{\boldsymbol{\rho}}_{j}^\top}{\widehat{\boldsymbol{\rho}}_{j}^\top \widehat{\boldsymbol{\rho}}_{j}} \widehat{\boldsymbol{\rho}}_j \right) \\
	&= P(\widehat{\boldsymbol{\rho}}_j-\widehat{\boldsymbol{\rho}}_j) \\
	&= \mathbf{0} \in \mathbb{R}^n, 
noting that $\int_{\mathcal{T}} \widetilde{W}^{[j]}[\mathbf{t}] \widehat{\xi}_j[\mathbf{t}] d\mathbf{t} = \widehat{\boldsymbol{\rho}}_j$. Now by the generalized eigenvalue problem (i.e., generalized Rayleigh quotient problem) presented in Proposition 2, we have that for $j < l$,
	\langle \langle \widehat{\xi}_l, \widehat{\xi}_j \rangle \rangle &= \widehat{\xi}_l^{*\top} (J^\ast + \Lambda \ddot{J}^\ast)\widehat{\xi}_j^{*} \\
	& = \frac{1}{\kappa_l} (V^\ast_l \widehat{\xi}_l^{*})^\top \widehat{\xi}_j^{*} \\
	&=  \frac{1}{\kappa_l}\widehat{\xi}_l^{* \top} V^\ast_l \widehat{\xi}_j^{*} \\
	&= \frac{1}{\kappa_l}\widehat{\xi}_l^{* \top} \left(n^{-2} J^\ast \widetilde{C}^{*[l]\top} \widetilde{\mathbf{y}}^{[l]}\widetilde{\mathbf{y}}^{[l]\top} \widetilde{C}^{*[l]} J^\ast \right) \widehat{\xi}_j^{*} \\
	&= \frac{1}{n^2\kappa_l}\widehat{\xi}_l^{* \top} \ J^\ast \widetilde{C}^{*[l]\top} \widetilde{\mathbf{y}}^{[l]}\widetilde{\mathbf{y}}^{[l]\top}  \widetilde{C}^{*[l]}\left(\int_{\mathcal{T}} B^*[\mathbf{t}] B^*[\mathbf{t}]^\top d\mathbf{t}  \right) \widehat{\xi}_j^{*} \\
	&= \frac{1}{n^2\kappa_l}\widehat{\xi}_l^{* \top}  J^\ast \widetilde{C}^{*[l]\top} \widetilde{\mathbf{y}}^{[l]}\widetilde{\mathbf{y}}^{[l]\top}  \left(\int_{\mathcal{T}} \widetilde{C}^{*[l]} B^*[\mathbf{t}] B^*[\mathbf{t}]^\top \widehat{\xi}_j^{*} d\mathbf{t}  \right)   \\
	&= \frac{1}{n^2\kappa_l}\widehat{\xi}_l^{* \top}  J^\ast \widetilde{C}^{*[l]\top} \widetilde{\mathbf{y}}^{[l]}\widetilde{\mathbf{y}}^{[l]\top}  \left(\int_{\mathcal{T}} \widetilde{W}^{[l]}[\mathbf{t}] \widehat{\xi}_j[\mathbf{t}] d\mathbf{t}  \right)   \\
	&=0, 
where the last equality follows from the above result that $\int_{\mathcal{T}} \widetilde{W}^{[l]}[\mathbf{t}] \widehat{\xi}_j[\mathbf{t}] d\mathbf{t} = \mathbf{0}$.

>>>>

\subsection{Preliminary step 1: finite-basis approximation}\label{sec: finite basis approximation}	
Let ${b_m(t)}$ be a twice-differentiable basis of $L^2([0,1])$ whose second derivatives are also linearly independent, for example, cubic B-splines, the Fourier basis, or an orthonormal polynomial basis of degree greater than three. Using this basis, 
the $j$th functional predictor, 
regression coefficient, PLS component direction, and orthogonalization regression coefficient (with iteration indices suppressed) are represented as follows:
\begin{equation*}
	X_{ij}(t) = \sum_{m=1}^\infty \theta_{ijm} \, b_m(t),~
	\beta_j(t) = \sum_{m=1}^\infty \eta_{jm} \, b_m(t),~
	%%%%%%%%%%%%%%%%%%%%%%
	\xi_j(t)
	=
	\sum_{m=1}^\infty
	\gamma_{jm}
	b_m(t),
	~
	\delta_j(t)
	=
	\sum_{m=1}^\infty
	\pi_{jm}
	b_m(t)		
\end{equation*}	
In practice, the full set of coefficients can not be obtained  with finite sample size, as functional data are measured on a finite grid. Thus, we truncate the expansion at $M$ terms and define the truncated basis space
$
\mathcal{B} := \operatorname{span}\bigl(b_1(t), \ldots, b_M(t)\bigr).
$
We choose a moderately large $M$ (e.g., 15 or 20) to capture functional variation without fine-tuning, as smoothness is handled via penalization (see Section~\ref{sec: Regularization of PLS Components}).
The truncated expansions of the predictor and coefficient are denoted as
\begin{equation*}
	\widetilde{X}_{ij}(t) := \sum_{m=1}^M \theta_{ijm} \, b_m(t), \quad
	\tilde{\beta}_j(t) := \sum_{m=1}^M \eta_{jm} \, b_m(t),
\end{equation*}
and our suggest method restricts each PLS component direction and orthogonalization regression coefficient to admit the following expansion:
\begin{equation}\label{def:truncated_zeta_and_delta}
	\xi_j(t) = \sum_{m=1}^M \gamma_{jm} \, b_m(t),
	\quad
	\delta_j(t)
	=
	\sum_{m=1}^M
	\pi_{jm}
	b_m(t)		
\end{equation}
This implies that all computations in this paper are carried out entirely within the subspace
\begin{equation}\label{def:truncated_hilbert_space}
	\widetilde{ \mathcal{H} } := \mathcal{B}^K \times \mathbb{R}^p \subset \mathcal{H}.
\end{equation}
The $i$th hybrid predictor, projected on $\widetilde{ \mathcal{H} }$, is represented by the tuple
\begin{equation*}
	\widetilde{W}_i := 
	(
	\widetilde{X}_{i1}, \ldots, \widetilde{X}_{iK}, \boldsymbol{Z}_i
	).
\end{equation*}
Let $\boldsymbol{\theta}_{ij}$, $\boldsymbol{\eta}_j$,  $\boldsymbol{\gamma}_j$, and
$\boldsymbol{\pi}_j$
denote the $M$-dimensional vectors of coefficients:
\begin{equation}\label{def:coef_vector_form}
	\boldsymbol{\theta}_{ij} := (\theta_{ij1}, \ldots, \theta_{ijM})^\top,
	~
	\boldsymbol{\eta}_{j} := (\eta_{j1}, \ldots, \eta_{jM})^\top ,~
	\boldsymbol{\gamma}_{j} := (\gamma_{j1}, \ldots, \gamma_{jM})^\top,~
	\boldsymbol{\pi}_{j} := (\pi_{j1}, \ldots, \pi_{jM})^\top
\end{equation}
For the predictors, we stack the coefficient vectors across observations into the matrix
\begin{equation}\label{def:theta}
	\Theta_j := (\boldsymbol{\theta}_{1j}, \ldots, \boldsymbol{\theta}_{nj})^\top  \in \mathbb{R}^{n \times M},
\end{equation}
and construct the full design matrix 
\begin{equation}\label{def:full_theta}
\Theta := (\Theta_1, \ldots, \Theta_K, Z) \in \mathbb{R}^{n \times (MK + p)}.
\end{equation}
Let us denote the response vector as $\mathbf{y} := (y_1, \ldots, y_n)^\top$.
Let $B, B^{\prime\prime} \in \mathbb{R}^{M \times M}$ denote the Gram matrices of the basis functions and their second derivatives, with entries
\begin{equation}\label{def:B_matrix}
	B_{m, m'} := \int_0^1 b_m(t)\, b_{m'}(t)\, dt, \quad
	B^{\prime\prime}_{m, m'} := \int_0^1 b_m''(t)\, b_{m'}''(t)\, dt,
\end{equation}
for $m, m' = 1, \ldots, M$. We then define the block-diagonal matrices
\begin{equation}\label{def:gram_block}
	\mathbb{B} := \operatorname{blkdiag}(B, \ldots, B, I_p), \quad
	\mathbb{B}^{\prime\prime} := \operatorname{blkdiag}(B^{\prime\prime}, \ldots, B^{\prime\prime}, I_p),
\end{equation}

@@

		\xi_j(t) \, \xi_j(t)
		\, dt

@@

\section{Proof of Lemma \ref{proposition:closed_form_orthgonalization} }\label{section:proof:proposition:closed_form_orthgonalization}
For notational simplicity, we omit the iteration superscript $[l]$.
 Recall from  \eqref{def:plsscore} that
$\hat{\rho}_i
=
\langle \widetilde{W}_i, \, \hat{\xi} \rangle_{\mathcal{H}}$.
Using the basis expansion coefficient notation from \eqref{def:truncated_zeta_and_delta}, \eqref{def:coef_vector_form}, \eqref{def:theta}, and \eqref{def:B_matrix}, the full vector of PLS scores $\hat{\boldsymbol{\rho}}  = (\hat{\rho}_1, \ldots, \hat{\rho}_n)^\top$ 
is computed through the following matrix multiplication:
\begin{equation}\label{score_inner_product_all_n}
	\hat{\boldsymbol{\rho}}^\top = \sum_{k=1}^K ( \hat{\boldsymbol{\gamma}}_j )^\top B \, \Theta_j^\top + \boldsymbol{\zeta}^\top \mathbf{Z}^\top.
\end{equation}
The criterion $\operatorname{PENSSE}( \delta)$ can be decomposed as follows:

@@

\begin{proof}
	The unit norm condition is trivially met by the constraint 
	$1 =  	\hat{\boldsymbol{\xi}}_l^\top 
	\hspace{-.2em}
	(\mathbb{B} + \Lambda \mathbb{B}^{\prime \prime}) \hat{\boldsymbol{\xi}}_l$
	enforced in \eqref{eq: Regularized generalized rayleigh quotient equation}, because we have:
	\begin{align*}
  	\hat{\boldsymbol{\xi}}_l^\top 
\hspace{-.2em}
(\mathbb{B} + \Lambda \mathbb{B}^{\prime \prime}) \hat{\boldsymbol{\xi}}_l
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&= \hat{\boldsymbol{\xi}}^\top_l 
\hspace{-.2em}
\mathbb{B}
 \hat{ \boldsymbol{\xi} }_l
 +
 \hat{ \boldsymbol{\xi} }_l^\top 
 \hspace{-.2em}
 \Lambda \mathbb{B}^{\prime \prime} \hat{ \boldsymbol{\xi} }_l
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \\&\overset{(i)}{=}
 \sum_{j=1}^K
 \int_0^1
 \hat{\xi}_l(t) \, \hat{\xi}_l(t)
 \, dt
 +
 \hat{\boldsymbol{\zeta}}^\top  \hat{\boldsymbol{\zeta}}
 +
  \hat{\boldsymbol{\xi}}^\top 
 \hspace{-.2em}
 \Lambda \mathbb{B}^{\prime \prime} \hat{\boldsymbol{\xi}}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \\&\overset{(ii)}{=}
 \sum_{j=1}^K
 \int_0^1
 \hat{\xi}_l(t) \, \hat{\xi}_l(t)
 \, dt
 +
 \hat{\boldsymbol{\zeta}}^\top \hat{\boldsymbol{\zeta}}
 +
 \sum \limits_{j=1}^K 
 \lambda_l
 \int_0^1 \bigl\{ \hat{\xi}_l^{\prime\prime}(t) \bigr\}^2 dt
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \\&=
 \langle \widehat{\xi}_{l}, \widehat{\xi}_{l} \rangle_{\mathcal{H}, \Lambda},
\numberthis \label{rough_norm_matrix_form}
	\end{align*}
where step $(i)$ uses \eqref{unit_norm_as_quadform}
and
step $(ii)$ uses \eqref{pen_as_matrix}.

Now to switch our gears toward the the orthogonality. For  $l_1 > l_2 $, we have
\begin{align*}
	\langle \widehat{\xi}_{l_1}, \widehat{\xi}_{l_2} \rangle_{\mathcal{H}, \Lambda}
	& \overset{(i)}{=}
	\hat{\boldsymbol{\xi}}_{l_1}^\top 
	\hspace{-.2em}
	(\mathbb{B} + \Lambda \mathbb{B}^{\prime \prime}) \hat{\boldsymbol{\xi}}_{l_2}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\	& \overset{(ii)}{=} \frac{1}{\kappa_{l_1}} (V^{[l_1]} \widehat{\boldsymbol{\xi}}_{l_1})^\top \hat{\boldsymbol{\xi}}_{l_2}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\	& = \frac{1}{\kappa_{l_1}}
	\widehat{\boldsymbol{\xi}}_{l_1}^\top
	V^{[l_1]} \hat{\boldsymbol{\xi}}_{l_2} 
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\	& \overset{(iii)}{=} \frac{1}{ n^2 \kappa_{l_1}}
	\widehat{\boldsymbol{\xi}}_{l_1}^\top
	(\mathbb{B} \Theta^{[l_1]\top} \mathbf{y})(\mathbb{B} \Theta^{[l_1]\top} \mathbf{y})^\top
	\hat{\boldsymbol{\xi}}_{l_2}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\	& = \frac{1}{ n^2 \kappa_{l_1}}
	\widehat{\boldsymbol{\xi}}_{l_1}^\top
	(\mathbb{B} \Theta^{[l_1]\top} \mathbf{y})
	\mathbf{y}^\top
	\Theta^{[l_1] } 
	\mathbb{B}
	\hat{\boldsymbol{\xi}}_{l_2}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\	& = c
	\mathbf{y}^\top
	(
	\Theta^{[l_1] }
	\mathbb{B}
	\hat{\boldsymbol{\xi}}_{l_2}
	),
\end{align*}
where
step $(i)$ uses \eqref{rough_norm_matrix_form},
step $(ii)$ uses the fact that $\hat{\boldsymbol{\xi}}_{l_1}$ is a generalized eigenvector ($\kappa_{l_1}$ denotes the corresponding generalized eigenvalue), as presented in Proposition \ref{proposition:eigen_regul},
and step $(iii)$ uses the definition of $V$ matrix given in \eqref{def:V_matrix},
and $c$ represents a scalar.

Our next task is to show that 
\begin{equation*}
\Theta^{[l_1] }
\mathbb{B}
\hat{\boldsymbol{\xi}}_{l_2} = \mathbf{0} \in \mathbb{R}^n.
\end{equation*}
Without loss of generality, we show that the $i$th entry of $\Theta^{[l_1] }
\mathbb{B}
\hat{\boldsymbol{\xi}}_{l_2}$ is 0.
From the construction in \eqref{def:full_theta}, \eqref{def:theta} and \eqref{def:gram_block}, the $i$-th entry is
\begin{align*}
	(\Theta^{[l_1] }
	\mathbb{B}
	\hat{\boldsymbol{\xi}}_{l_2})_i
	&=
	(\theta_{i11}^{[l_1]}, \ldots, \theta_{i1M}^{[l_1]}, \ldots, \theta_{iK1}^{[l_1]}, \ldots, \theta_{iKM}^{[l_1]}, Z_{i1}^{[l_1]}, \ldots, Z_{ip}^{[l_1]})
	\mathbb{B}
	\hat{\boldsymbol{\xi}}_{l_2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\&=
	\sum_{k=1}^K
	\boldsymbol{\theta}_{ik}^{[l_1]} B
	\hat{\boldsymbol{\gamma}}_k^{[l_2]}
	+
	\mathbf{Z}_{i}^{[l_1] \top } \hat{\boldsymbol{\zeta}}^{[l_2]}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\\	&=
\langle
\widetilde{W}_i^{[l_1]}, 
\hat{\xi}^{[l_2]}
\rangle_{\mathcal{H}},
\end{align*}
where the last equality uses the computation of the $i$th entry in \eqref{score_inner_product_all_n}.

We can express the $l_1$-th residualized hybrid predictor in terms of its previous $(l_1-1)$-th version:
\begin{equation*}
	\widetilde{W}_i^{[l_1]}= \widetilde{W}_i^{[l_1-1]} - \widehat{{\rho}}_i^{[l_1-1]} \widehat{\delta}^{[l_1-1]}
\end{equation*}
\end{proof}

which, after applying the finite-dimensional form (9), implies that
\begin{align*}
	\widetilde{C}^{*[l]}B^*[\mathbf{t}] &=\widetilde{C}^{*[l-1]}B^*[\mathbf{t}] - \widehat{\boldsymbol{\rho}}_{l-1}\widehat{\boldsymbol{\gamma}}^*_{l-1}B^*[\mathbf{t}] \\
	&= \widetilde{C}^{*[l-1]}B^*[\mathbf{t}] - \widehat{\boldsymbol{\rho}}_{l-1} \widehat{\boldsymbol{\rho}}_{l-1}^\top \widetilde{C}^{*[l-1]} J^\ast \left(\widehat{\boldsymbol{\rho}}_{i,l-1}^\top \widehat{\boldsymbol{\rho}}_{l-1} J^\ast \right)^{-1} B^*[\mathbf{t}] \\
	& =\widetilde{C}^{*[l-1]}B^*[\mathbf{t}] -  \left(\frac{\widehat{\boldsymbol{\rho}}_{l-1} \widehat{\boldsymbol{\rho}}_{l-1}^\top}{\widehat{\boldsymbol{\rho}}_{l-1}^\top \widehat{\boldsymbol{\rho}}_{l-1}} \right) \widetilde{C}^{*[l-1]} B^*[\mathbf{t}] \\
	&= \left(I_n - \frac{\widehat{\boldsymbol{\rho}}_{l-1} \widehat{\boldsymbol{\rho}}_{l-1}^\top}{\widehat{\boldsymbol{\rho}}_{l-1}^\top \widehat{\boldsymbol{\rho}}_{l-1}} \right) \widetilde{C}^{*[l-1]} B^*[\mathbf{t}].
\end{align*}

	Let us denote
$
\hat{\boldsymbol{\rho}}^{[l]} := (\hat{\rho}^{[l]}_1, \ldots, \hat{\rho}^{[l]}_n)^\top.
%\quad\text{and} \quad		\bar{\rho}^{[l]} := (\hat{\boldsymbol{\rho}}^{[l]})^\top \hat{\boldsymbol{\rho}}^{[l]}.
$
The solution to the optimization problem \eqref{def:argmin_pensse}, denoted
$\hat{\delta}^{[l]}   \in \tilde{\mathcal{H}}$, has basis coefficients
\begin{equation*}
	\hat{\boldsymbol{\pi}}_j^{[l]}
	=
	\bigl(
	\| 	\hat{\boldsymbol{\rho}}^{[l]}\|_2^2
	B
	+
	\tau
	B^{\prime \prime}
	\bigr)^{-1}
	B
	(\Theta_j^{[l]})^\top
	\hat{\boldsymbol{\rho}}^{[l]}
	,
	~
	j = 1, \ldots, K,
\end{equation*}
and scalar parts given by:
\begin{equation*}
	\hat{\boldsymbol{\chi} }^{[l]}
	=
	(\mathbf{Z}^{[l]})^\top  \hat{\boldsymbol{\rho}}^{[l]}.
\end{equation*}
